---
title: "Packages"
subtitle: "**Celtic Sea Sprat  (Sprattus sprattus)**"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike - 0 International Public License
#knit: (function(inputFile, encoding) {
#  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
#bibliography: refs.bib
editor_options: 
  markdown: 
    wrap: 72
---


```{r, knitr, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =TRUE, 
               eval      =TRUE,
               
               cache     =TRUE, 
               cache.path="cache/00/",
               
               fig.path  ="../tex/figs/00-",
               fig.width =8,
               fig.height=6,
               dev       ="png")

options(digits=3)

iFig=0
iTab=0
```



# R Package Descriptions

## Libraries

A variety of R packages are required for data manipulation and plotting,
obtaining life history parameters, and modelling.


This document provides descriptions for various R packages, organsed by application.

### Data Manipulation and Plotting

These libraries are part of the R ecosystem for data manipulation and visualization, each serving specific purposes from data restructuring (plyr, dplyr, reshape) to advanced plotting (ggplot2, GGally, ggpubr, ggcorrplot).


```{r, pkgs-ggplot}
library(plyr)
library(dplyr)
library(reshape)
library(ggplot2)

library(ggplot2); theme_set(theme_bw())
library(GGally)
library(ggpubr)
library(ggcorrplot)
```


- **plyr**: A set of tools for splitting, applying, and combining data. It's designed to work with data frames, lists, and arrays, making it easier to manipulate data structures in R. [Vignette](https://www.rdocumentation.org/packages/plyr/versions/- 8.9)[10][13][15][17][19].
   
- **dplyr**: A grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges. It's focused on tools for working with data frames. [Vignette](https://dplyr.tidyverse.org/articles/dplyr.html)[2][5][11][14][16][18][20].

- **reshape**: Allows for flexibly restructuring and aggregating data using just two functions: melt and cast. It simplifies the process of transforming data between wide and long formats. [Documentation](https://cran.r-project.org/web/packages/reshape/index.html)[3][9].

- **ggplot2**: A system for declaratively creating graphics, based on The Grammar of Graphics. It allows for the creation of complex plots from data in a data frame. [Vignette](https://ggplot- tidyverse.org)[4].

- **ggpubr**: Provides easy-to-use functions for creating and customizing 'ggplot2'-based publication-ready plots. It simplifies the process of creating common types of plots and adjusting their appearance. [Vignette](https://rpkgs.datanovia.com/ggpubr/)[7].

- **ggcorrplot**: A visualization tool for displaying a correlation matrix using ggplot-  It simplifies the process of creating visually appealing correlation matrices. [Vignette](https://cran.r-project.org/web/packages/ggcorrplot/vignettes/ggcorrplot-intro.html)[7].

- **GGally**: Extends ggplot2 by adding several functions to reduce the complexity of combining geometric objects with transformed data. It's useful for creating pairs plots, correlations, and other complex relationships. [Vignette](https://ggobi.github.io/ggally/)[7].



Citations:
[1] https://cran.r-project.org/web/packages/plyr/index.html
[2] https://cran.r-project.org/web/packages/dplyr/index.html
[3] https://cran.r-project.org/web/packages/reshape/reshape.pdf
[4] https://www.rdocumentation.org/packages/plyr/versions/- 8.9
[5] https://cran.r-project.org/web/packages/dplyr/dplyr.pdf
[6] https://www.statmethods.net/management/reshape.html
[7] https://ademos.people.uic.edu/Chapter- html
[8] https://datacarpentry.org/R-genomics/04-dplyr.html
[9] https://cran.r-project.org/web/packages/reshape/index.html
[10] https://rdrr.io/cran/plyr/
[11] https://dplyr.tidyverse.org/articles/dplyr.html
[12] https://seananderson.ca/2013/10/19/reshape/
[13] https://github.com/hadley/plyr
[14] https://www.geeksforgeeks.org/dplyr-package-in-r-programming/
[15] https://cran.r-project.org/web/packages/plyr/plyr.pdf
[16] https://www.rdocumentation.org/packages/dplyr/versions/0.- 0
[17] https://www.rdocumentation.org/packages/plyr/versions/- 8.9/topics/plyr
[18] https://github.com/tidyverse/dplyr/blob/main/DESCRIPTION
[19] https://seananderson.ca/courses/12-plyr/plyr_201- pdf
[20] https://bookdown.org/ndphillips/YaRrr/dplyr.html

### FLR (Fisheries Library in R)

```{r, pkgs-flr}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(ggplotFL)
```

-  **FLCore**: Provides the core classes and methods to represent typical fisheries science data types and operations on them. It's the backbone of the FLR project for quantitative fisheries science. [Vignette](https://flr-project.org/FLCore/).

-  **FLBRP**: Stands for FL Biological Reference Points. It's a package in the FLR project for estimating biological reference points from stock assessment models. [Documentation](https://flr-project.org/packages/).

-  **FLasher**: A package for implementing stock assessment models, designed to be flexible and user-friendly for fisheries scientists. [Vignette](https://flr-project.org/FLasher/).

-  **FLife**: Provides life history models for fish populations as part of the FLR project. It's used for studying and simulating fish population dynamics. [Documentation](https://flr-project.org/packages/).

-  **ggplotFL**: Extends ggplot2 for plotting FLR objects. It allows for the creation of publication-quality plots of fisheries science data represented using FLR classes. [Vignette](https://flr-project.org/ggplotFL/).

### Parallel Computing

#### Parallel Computing

```{r, pkgs-paral}
library(doParallel)
library(foreach)
```


-  **doParallel**: Provides a parallel backend for the 'foreach' looping construct. It allows for easy parallelization of R code on platforms supporting the 'parallel' package. [Vignette](https://cran.r-project.org/web/packages/doParallel/vignettes/gettingstartedParallel.pdf).

-  **foreach**: Provides a looping construct for executing R code repeatedly. 'foreach' can be used with parallel backends provided by other packages like 'doParallel' for parallel execution. [Vignette](https://cran.r-project.org/web/packages/foreach/vignettes/foreach.pdf).


#### Life Histories

```{r, pkgs-lh}
library(rfishbase)
library(SPMpriors)
library(FishLife)
```


-  **rfishbase**: A programmatic interface to 'FishBase', the world's largest database of fish species. It allows for easy access to a wide range of fish species data in R. [Vignette](https://cran.r-project.org/web/packages/rfishbase/vignettes/intro.html).

-  **SPMpriors**: Provides a way to generate prior distributions for stock productivity parameters, useful in stock assessment models and fisheries management. [Documentation](https://cran.r-project.org/web/packages/SPMpriors/index.html) (No direct vignette link, see package documentation).

-  **FishLife**: Offers predictions of life history, morphological, ecological, and phylogenetic traits of fish based on the 'FishLife' database. It's useful for ecological and evolutionary research. [Vignette](https://cran.r-project.org/web/packages/FishLife/vignettes/FishLife.html).

Please note that for some packages, a direct link to a vignette is not provided, but you can find more information and vignettes (if available) on their respective CRAN pages or package documentation.


## Parallel Computing

### doParallel
Provides a parallel backend for the 'foreach' looping construct, enabling the execution of R code in parallel on multiple cores or machines. [Vignette](https://cran.r-project.org/web/packages/doParallel/index.html).

### foreach
Implements a looping construct for R, which is designed to be used in parallel computing environments. It's flexible and allows for the execution of R code in parallel. [Vignette](https://cran.r-project.org/web/packages/foreach/index.html).

## Life Histories

### rfishbase
A programmatic interface to 'FishBase', the world's largest database of fish species. It allows for easy access to a wide range of data on fish species. [Vignette](https://cran.r-project.org/web/packages/rfishbase/index.html).

### SPMpriors
Provides a way to generate prior distributions for stock productivity parameters, useful in fisheries stock assessment models. [Documentation](https://cran.r-project.org/web/packages/SPMpriors/index.html).

### FishLife
Offers predictions of life-history traits for fish species, based on phylogenetic imputation. It's useful for filling gaps in life-history data in fishery stock assessments. [Vignette](https://cran.r-project.org/web/packages/FishLife/index.html).


#### Statistics

```{r, pkgs-stat}
library(statcomp)
#library(RTseries)
```

### statcomp
Facilitates statistical complexity and entropy-based analysis of time series data. It includes various methods for chaos, fractal, and complexity analysis. [Vignette](https://cran.r-project.org/web/packages/statcomp/index.html).

### RTseries
Provides tools for analyzing and manipulating time series data, including functions for smoothing, decomposition, and forecasting. [Documentation](https://cran.r-project.org/web/packages/RTseries/index.html).

#### Fisheries

```{r, pkgs-fish}
library(FSA)
library(fishmethods)
```

### FSA
Offers functions for fishery science methods and models, including age-length keys, stock-recruitment relationships, and catch-at-age analysis. [Vignette](https://cran.r-project.org/web/packages/FSA/index.html).

### fishmethods
Contains a wide range of methods and models used in fishery science, such as mortality estimation, stock assessment, and length-frequency analysis. [Vignette](https://cran.r-project.org/web/packages/fishmethods/index.html).

## Utilities


#### Utilities

```{r, pkgs-tools}
library(broom)
library(bootstrap)
library(nlstools)
```

### broom
Simplifies the process of converting statistical analysis objects in R into tidy data frames, making it easier to work with model outputs. [Vignette](https://cran.r-project.org/web/packages/broom/index.html).

### bootstrap
Provides functions for bootstrapping, which is a method for estimating the distribution of a statistic by resampling with replacement from the data. [Documentation](https://cran.r-project.org/web/packages/bootstrap/index.html).

### nlstools
Offers tools for analyzing nonlinear regression models, including diagnostic plots and confidence interval estimation. [Vignette](https://cran.r-project.org/web/packages/nlstools/index.html).


### Installation 

First the required Libraries (i.e. packages) need to be installed. This can be be done
from cran, e.g.

Set mirror if needed `chooseCRANmirror()`

```{r, install-util, eval=FALSE}
install.packages(c("plyr", 
                   "dplyr", 
                   "reshape"))
```

```{r, install-gg, eval=FALSE}
install.packages(c("ggplot2",
                   "ggpubr",
                   "ggcorrplot",
                   "GGally"))
```

#### FLR

Or from <https://flr-project.org/>

To install the latest released versions of the FLR packages, and their dependencies
enter

```{r, eval=FALSE}
source("http://flr-project.org/R/instFLR.R")
```

then select 'FLCore', 'ggplotFL', 'FLBRP', 'FLasher', 'FLife', 'mydas'

Or install from source

```{r, eval=FALSE}
install.packages("remotes")

library(remotes)

remotes:::install_github("flr/FLCore")
remotes:::install_github("flr/FLBRP")
remotes:::install_github("flr/ggplotFL")
remotes:::install_github("flr/FLfishery")
remotes:::install_github("flr/FLasher", dependencies=TRUE)
remotes:::install_github("flr/FLife")
remotes:::install_github("flr/mydas")
```


#### FishBase

There are compatability problems with `FishBase`, so install an earlier
version. 

These have been fixed now. see <https://james-thorson-noaa.github.io/FishLife/>

```{r, eval=FALSE}
remotes::install_github("ropensci/rfishbase",ref="d150f2e0f5")
remotes::install_github("james-thorson/FishLife")
remotes::install_github("Henning-Winker/SPMpriors")
```

[Back to Top](#top)

## Examples {#example}

The examples here use `FLR`, tutorials on the various methods and
packages can be found at <https://flr-project.org/doc/>

The Fisheries Library in R (FLR) is a collection of tools for
quantitative fisheries science, developed in the R language, that
facilitates the construction of bio-economic simulation models of
fisheries systems. FLR builds on the powerful R environment and syntax
to create a domain-specific language for the quantitative analysis of
the expected risks and effects of fisheries management decision. The
classes and methods in FLR consider uncertainty an integral part of our
knowledge of fisheries system.

Life history traits include growth rate; age and size at sexual
maturity; the temporal pattern or schedule of reproduction; the number,
size, and sex ratio of offspring; the distribution of intrinsic or
extrinsic mortality rates (e.g., patterns of senescence); and patterns
of dormancy and dispersal. These traits contribute directly to
age-specific survival and reproductive functions. The
[`FLife`](https://flr-project.org/doc/Using_information_on_life_history_relationships.html)
package has a variety of methods for modelling life history traits and
functional forms for processes for use in fish stock assessment and for
conducting Management Strategy Evaluation.

These relationships have many uses, for example in age-structured
population models, functional relationships for these processes allow
the calculation of the population growth rate and have been used to to
develop priors in stock assesments and to parameterise ecological
models. The `FLife` package has methods for modelling functional forms,
for simulating equilibrium `FLBRP` and dynamic stock objects `FLStock`.

Create an `FLPar` object For a fish of length 45 cm.

```{r}
par=FLPar(linf=45)
```

```{r}
par
```

`FLPar()` creates an FLR object of
[class](https://flr-project.org/doc/An_overview_of_the_FLCore_classes.html)
`FLPar`

```{r}
is(par)
```

```{r}
?FLPar
```

Life history theory means that traits like length at maturity ($l_{50}$)
can be derived from the expected size of a species, the `lhPar` method
therefore uses life history theory to derive missing parameters.

```{r}
par=lhPar(par)
par
```

```{r}
?lhPar
```

An `FLBRP` object can then be created using `lhEql`

```{r}
eql=lhEql(par)
```

```{r}
?lhEql
```

This creates an object of class `FLBRP`

```{r}
is(eql)
```

[`FLBRP`](https://flr-project.org/doc/Reference_points_for_fisheries_management_with_FLBRP.html)
is used to estimate reference points, and model equilibrium dynamics

```{r}
refpts(eql)
```

[ggplotFL](https://flr-project.org/doc/ggplotFL_plotting_FLR_objects_with_ggplot- html)
is used for plotting FLR objects with ggplot2

```{r}
plot(eql)
```

```{r}
plot(eql,refpts="msy")
```

## More Information {#more}

-   For more information on the FLR Project for Quantitative Fisheries
    Science in R, visit the FLR webpage [^1].

[^1]: <http://flr-project.org>

## Author information

**Laurence Kell**.
[laurie\@seaplusplus.co.uk](mailto:laurie@seaplusplus.co.uk){.email}

## Acknowledgements

## Software Versions

**`r version$version.string`**

-   knitr: `r packageVersion('knitr')`

-   FLCore: `r packageVersion('FLCore')`

-   FLBRP: `r packageVersion('FLBRP')`

-   FLasher: `r packageVersion('FLasher')`

-   FLife: `r packageVersion('FLife')`

-   ggplotFL: `r packageVersion('ggplotFL')`

-   rfishbase: `r packageVersion('rfishbase')`

-   SPMpriors: `r packageVersion('SPMpriors')`

-   FishLife: `r packageVersion('FishLife')`

-   plyr: `r packageVersion('plyr')`

-   dplyr: `r packageVersion('dplyr')`

-   reshape: `r packageVersion('reshape')`

-   ggplot2: `r packageVersion('ggplot2')`

-   ggpubr: `r packageVersion('ggpubr')`

-   ggcorrplot: `r packageVersion('ggcorrplot')`

-   GGally: `r packageVersion('GGally')`

**Compiled**: `r date()`

[Back to Top](#top)

[Back to Top](#top)

# References {#references}

[Back to Top](#top)
