---
title: "Indicator Skill"
subtitle: "**Celtic Sea Sprat  (Sprattus sprattus)**"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#bibliography: /home/laurie/Desktop/refs.bib
---
  
```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="cache/03_skill/",
               fig.path  ="../figs/03_skill/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```


```{r, pkgs}
library(ggplot2)

library(FLCore)
library(FLBRP)
library(FLFishery)
library(FLasher)
library(FLife)

library(ggpubr)
library(GGally)
library(ggcorrplot)

library(plyr)
library(dplyr)
library(reshape)

library(SPMpriors)
library(FishLife)
```

```{r, oms}
load("../data/om/omScenarios.RData")
load("../data/om/devs.RData")
```

```{r, source}
source("../source/roc.R")
```


```{r}
targetF=0.1
```


```{r om-plot, fig.height = 6, fig.width = 8}
p1=plot(window(FLStocks("Status Quo"=burnin[[1]],"Kicked"=kick[[1]],
                        "No Demand"=nodemand[[1]],"F Trend"=ftrend[[1]]),start=2010),metrics=list(
    Rec  =function(x) rec(x)[,,,3],
    SSB  =function(x) ssb(x)[,,,2],
    Catch=function(x) apply(catch(x),c(2,6), sum),
    F     =function(x) apply(fbar(x), c(2,6),mean),
    Forage=function(x) apply(stock.wt(x)%*%(m(x)-0.1)/(z(x))%*%(1-exp(-z(x))),c(2,6),sum)),
    iter=c(1,15))+
  geom_flpar(data=FLPars(Rec  =FLPar("Rmsy"=refpts(eq4)[c("msy"),"rec",1,drop=T])*0.6,
                           SSB  =FLPar("Bmsy"=refpts(eq4)[c("msy"),"ssb",1,drop=T]),
                           F    =FLPar("Fmsy"=refpts(eq4)[c("msy"),"harvest",1,drop=T])/4,
                           Catch=FLPar("MSY" =refpts(eq4)[c("msy"),"yield",1,drop=T])*4),x=rep(ISOdate(2012,1,1),4))+
    theme_bw()+xlab("Year")+theme(legend.position="bottom")
```

```{r om-plot-2, fig.height = 6, fig.width = 8}
p2=plot(window(FLStocks("Status Quo"=burnin[[2]],"Kicked"=kick[[2]],
                        "No Demand"=nodemand[[2]],"F Trend"=ftrend[[2]]),start=2010),metrics=list(
    Rec  =function(x) rec(x)[,,,3],
    SSB  =function(x) ssb(x)[,,,2],
    Catch=function(x) apply(catch(x),c(2,6), sum),
    F     =function(x) apply(fbar(x), c(2,6),mean),
    Forage=function(x) apply(stock.wt(x)%*%(m(x)-0.1)/(z(x))%*%(1-exp(-z(x))),c(2,6),sum)),
    iter=c(1,15))+    geom_flpar(data=FLPars(Rec  =FLPar("Rmsy"=refpts(eq4)[c("msy"),"rec",1,drop=T])*0.6,
                           SSB  =FLPar("Bmsy"=refpts(eq4)[c("msy"),"ssb",1,drop=T]),
                           F    =FLPar("Fmsy"=refpts(eq4)[c("msy"),"harvest",1,drop=T])/4,
                           Catch=FLPar("MSY" =refpts(eq4)[c("msy"),"yield",1,drop=T])*4),x=rep(ISOdate(2012,1,1),4))+
    theme_bw()+xlab("Year")+theme(legend.position="bottom")
```

```{r om-plot-3, fig.height = 6, fig.width = 8}
p3=plot(window(FLStocks("Status Quo"=burnin[[3]],"Kicked"=kick[[3]],
                        "No Demand"=nodemand[[3]],"F Trend"=ftrend[[3]]),start=2010),metrics=list(
    Rec  =function(x) rec(x)[,,,3],
    SSB  =function(x) ssb(x)[,,,2],
    Catch=function(x) apply(catch(x),c(2,6), sum),
    F     =function(x) apply(fbar(x), c(2,6),mean),
    Forage=function(x) apply(stock.wt(x)%*%(m(x)-0.1)/(z(x))%*%(1-exp(-z(x))),c(2,6),sum)),
    iter=c(1,15))+    geom_flpar(data=FLPars(Rec  =FLPar("Rmsy"=refpts(eq4)[c("msy"),"rec",1,drop=T])*0.6,
                           SSB  =FLPar("Bmsy"=refpts(eq4)[c("msy"),"ssb",1,drop=T]),
                           F    =FLPar("Fmsy"=refpts(eq4)[c("msy"),"harvest",1,drop=T])/4,
                           Catch=FLPar("MSY" =refpts(eq4)[c("msy"),"yield",1,drop=T])*4),x=rep(ISOdate(2012,1,1),4))+
    theme_bw()+xlab("Year")+theme(legend.position="bottom") 
```


```{r om-plot-4, fig.height = 6, fig.width = 8}
p4=plot(window(FLStocks("Status Quo"=burnin[[4]],"Kicked"=kick[[4]],
                        "No Demand"=nodemand[[4]],"F Trend"=ftrend[[4]]),start=2010),metrics=list(
    Rec  =function(x) rec(x)[,,,3],
    SSB  =function(x) ssb(x)[,,,2],
    Catch=function(x) apply(catch(x),c(2,6), sum),
    F     =function(x) apply(fbar(x), c(2,6),mean),
    Forage=function(x) apply(stock.wt(x)%*%(m(x)-0.1)/(z(x))%*%(1-exp(-z(x))),c(2,6),sum)),
    iter=c(1,15))+  
    geom_flpar(data=FLPars(Rec  =FLPar("Rmsy"=refpts(eq4)[c("msy"),"rec",1,drop=T])*0.6,
                           SSB  =FLPar("Bmsy"=refpts(eq4)[c("msy"),"ssb",1,drop=T]),
                           F    =FLPar("Fmsy"=refpts(eq4)[c("msy"),"harvest",1,drop=T])/4,
                           Catch=FLPar("MSY" =refpts(eq4)[c("msy"),"yield",1,drop=T])*4),x=rep(ISOdate(2012,1,1),4))+
    theme_bw()+xlab("Year")+theme(legend.position="bottom")
```


```{r, plot-oms, fig.width=10, fig.height=6}
p1=p1+theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank() ) 
p2=p2+theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank()) 
p3=p3+theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank()) 
p4=p4+theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank())
            
ggarrange(p1,p2,p3,p4,label.y=0, 
          widths = c(2,2,2,2), heights = c(1),
          nrow   = 1,          ncol    = 4,
          labels = c("A","B"),
          common.legend = TRUE,
          align = "hv")+theme_bw()+
          theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Operating Model


## Abundance Index

```{r hr}
oms=FLStocks("Constant"=burnin[[1]],"No Demand"=nodemand[[1]],"Kicked"=kick[[1]],"F Trend"=ftrend[[1]])
hDevs=    rlnoise(100,stock(oms[[1]])[,,,3,,1]%=%0,0.3,0)%*%
     qmin(rlnoise(100,stock(oms[[1]])[,,,3,,1]%=%0,0.5,0.6),1)

mp=as.data.frame(FLQuants(llply(oms,function(x) hDevs%*%catch(x)[,ac(2031),,"3"]%/%biomass(x)[,ac(2031),,"3"])),drop=T)
om=as.data.frame(FLQuants(llply(oms,function(x) fbar(x)[, ac(2031),,"3"])),drop=T)
hr=merge(om,mp,by=c("iter","qname"))
```

```{r, fhDEvs, ig.height=4, fig.width=8}
plot(hDevs) 
```

**Figure `r iFig=iFig+1; iFig`** Survey devs

```{r, hDevs}
hDevs=hDevs[,dimnames(biomass(oms[[2]]))$year]
x=ssb(oms[[2]])[,,,2] 
y=hDevs%*%biomass(oms[[2]])[,,,"3"]

plot(FLQuants("OM"=x,"Index"=y))
```

**Figure `r iFig=iFig+1; iFig`** Index

```{r}
ggplot(model.frame(FLQuants("OM"=x,"Index"=y)))+
  geom_point(aes(OM,Index))
  #geom_vline(aes(xintercept=1),col="red")+
  #geom_hline(aes(yintercept=1),col="red")  
```

**Figure `r iFig=iFig+1; iFig`** Index v OM


## Harvest Rate

```{r hr-index, fig.height=6, fig.width=8}
plot(FLQuants("OM"=fbar(oms[[2]])[,,,"3"],
              "Index"=hDevs%*%catch(oms[[2]])[,,,"3"]%/%biomass(oms[[2]])[,,,"3"]))
```

**Figure `r iFig=iFig+1; iFig`** HR Index

```{r hr-plot, fig.height = 6, fig.width=6}
ggplot(hr)+  
  geom_point(aes(data.x,data.y))+
  xlab("OM")+ylab("Index")+
  geom_hline(aes(yintercept=0.025),col="red")+
  geom_vline(aes(xintercept=targetF),col="red")
```

**Figure `r iFig=iFig+1; iFig`** Comparison between $Catch/SSB$ and OM F.


```{r hr-roc}
rtn=with(hr,roc(data.x/targetF,data.y))

hDevs=qmin(rlnoise(100,stock(oms[[1]])[,,,3,,1]%=%0,0.3,0.5),1)

mp=as.data.frame(FLQuants(llply(oms,function(x) hDevs%*%catch(x)[,ac(2030),,"3"]%/%biomass(x)[,ac(2030),,"3"])),drop=T)
om=as.data.frame(FLQuants(llply(oms,function(x) fbar(x)[, ac(2031),,"3"])),drop=T)
hr=merge(om,mp,by=c("iter","qname"))

rtn2=with(hr,roc(data.x/targetF,data.y))

ggplot(rbind(cbind(Lag="1 Year",rtn2),
             cbind(Lag="Noner",rtn)))+
  geom_line(aes(FPR,TPR,col=Lag))
  geom_abline(aes(intercept=0,slope=1),col="red")
```

**Figure `r iFig=iFig+1; iFig`** ROC Curves Shifted by a year

## Length Based Indicators

```{r lbi}
oms=FLStocks("Constant"=burnin[[1]],"No Demand"=nodemand[[1]],"Kicked"=kick[[1]])
lbi=FLQuants(mlply(names(oms), function(qname) 
  llply(indicators.len(oms[[qname]][,ac(seq(2018,2025,1)),,"3"], indicators=c('lbar'), params=pars[,1], metric='catch.n'))[[1]]))
names(lbi)=names(oms)

oms=FLStocks("Constant"=burnin[[2]],"No Demand"=nodemand[[2]],"Kicked"=kick[[2]])
lbi2=FLQuants(mlply(names(oms), function(qname) 
  llply(indicators.len(oms[[qname]][,ac(seq(2018,2025,1)),,"3"], indicators=c('lbar'), params=pars[,1], metric='catch.n'))[[1]]))
names(lbi2)=names(oms)

oms=FLStocks("Constant"=burnin[[3]],"No Demand"=nodemand[[3]],"Kicked"=kick[[3]])
lbi3=FLQuants(mlply(names(oms), function(qname) 
  llply(indicators.len(oms[[qname]][,ac(seq(2018,2025,1)),,"3"], indicators=c('lbar'), params=pars[,1], metric='catch.n'))[[1]]))
names(lbi3)=names(oms)

oms=FLStocks("Constant"=burnin[[4]],"No Demand"=nodemand[[4]],"Kicked"=kick[[4]])
lbi4=FLQuants(mlply(names(oms), function(qname) 
  llply(indicators.len(oms[[qname]][,ac(seq(2018,2025,1)),,"3"], indicators=c('lbar'), params=pars[,1], metric='catch.n'))[[1]]))
names(lbi4)=names(oms)
```

```{r, lbi-plot, fig.height = 6, fig.width = 8}
save(lbi,file="tmp.Data")
load(         "tmp.Data")
names(lbi)=gsub(" ","_",names(lbi))
```

**Figure `r iFig=iFig+1; iFig`** Length based indicator

```{r lbi-2}
oms=FLStocks("Constant"=burnin[[1]],"No Demand"=nodemand[[1]],"Kicked"=kick[[1]])
mp=as.data.frame(FLQuants(llply(oms,function(x) hDevs%*%catch(x)[,ac(2031),,"3"]%/%biomass(x)[,ac(2031),,"3"])),drop=T)
om=as.data.frame(FLQuants(llply(oms,function(x) fbar(x)[, ac(2031),,"3"])),drop=T)
hr=merge(om,mp,by=c("iter","qname"))
lbi=transform(ldply(lbi, function(qname) as.data.frame(qname,drop=T)),qname=.id)[,-1]
lbi=merge(om,lbi,by=c("iter","qname"),all.y=T)  

oms=FLStocks("Constant"=burnin[[2]],"No Demand"=nodemand[[2]],"Kicked"=kick[[2]])
mp2=as.data.frame(FLQuants(llply(oms,function(x) hDevs%*%catch(x)[,ac(2031),,"3"]%/%biomass(x)[,ac(2031),,"3"])),drop=T)
om2=as.data.frame(FLQuants(llply(oms,function(x) fbar(x)[, ac(2031),,"3"])),drop=T)
hr2=merge(om2,mp2,by=c("iter","qname"))
lbi2=transform(ldply(lbi2, function(qname) as.data.frame(qname,drop=T)),qname=.id)[,-1]
lbi2=merge(om,lbi2,by=c("iter","qname"),all.y=T)  

oms=FLStocks("Constant"=burnin[[3]],"No Demand"=nodemand[[3]],"Kicked"=kick[[3]])
mp3=as.data.frame(FLQuants(llply(oms,function(x) hDevs%*%catch(x)[,ac(2031),,"3"]%/%biomass(x)[,ac(2031),,"3"])),drop=T)
om3=as.data.frame(FLQuants(llply(oms,function(x) fbar(x)[, ac(2031),,"3"])),drop=T)
hr3=merge(om3,mp3,by=c("iter","qname"))
lbi3=transform(ldply(lbi3, function(qname) as.data.frame(qname,drop=T)),qname=.id)[,-1]
lbi3=merge(om,lbi3,by=c("iter","qname"),all.y=T)  

oms=FLStocks("Constant"=burnin[[4]],"No Demand"=nodemand[[4]],"Kicked"=kick[[4]])
mp4=as.data.frame(FLQuants(llply(oms,function(x) hDevs%*%catch(x)[,ac(2031),,"3"]%/%biomass(x)[,ac(2031),,"3"])),drop=T)
om4=as.data.frame(FLQuants(llply(oms,function(x) fbar(x)[, ac(2031),,"3"])),drop=T)
hr4=merge(om4,mp4,by=c("iter","qname"))
lbi4=transform(ldply(lbi4, function(qname) as.data.frame(qname,drop=T)),qname=.id)[,-1]
lbi4=merge(om,lbi4,by=c("iter","qname"),all.y=T)
```

```{r lbi-2-plot, fig.height=6, fig.width=8}
ggplot(subset(lbi4,year>=2020))+  
  geom_point(aes(data.x,1/(data.y/c(pars["linf",1]*0.8)),col=qname))+
  xlab("OM")+ylab("Index")+
  geom_hline(aes(yintercept=1),col="red")+
  geom_vline(aes(xintercept=targetF),col="red")+
  facet_wrap(~year)+
  theme(legend.position="bottom") 
```

**Figure `r iFig=iFig+1; iFig`** Comparison of LBI with OM F.


```{r TSS}
rtn=ddply(subset(lbi, year>2020), .(year), with, roc(data.x/targetF,1/data.y))
rt2=ddply(subset(lbi2,year>2020), .(year), with, roc(data.x/targetF,1/data.y))
rt3=ddply(subset(lbi3,year>2020), .(year), with, roc(data.x/targetF,1/data.y))
rt4=ddply(subset(lbi4,year>2020), .(year), with, roc(data.x/targetF,1/data.y))

rtn=rbind(cbind(Scenario="Base",    rtn),
          cbind(Scenario="M Cohort",rt2),
          cbind(Scenario="M1",      rt3),
          cbind(Scenario="M2",      rt4)) 

#ddply(rtn, .(Scenario,year), with, transform(rtn[abs(rtn$ind-1)==min(abs(rtn$ind-1)),],roc.TSS=TPR-FPR)) 
```


```{r roc, fig.height = 6, fig.width = 8}
ggplot(rtn)+ 
  geom_line(aes(FPR,TPR,col=year,group=year))+geom_abline(aes(intercept=0,slope=1),col="red")+
  facet_wrap(~Scenario)
```

**Figure `r iFig=iFig+1; iFig`** ROC Curves


## Trends


## Two over three rule

```{r}
#fn=function(x,f=2,t=3) 
# sum(x[length(x)-(0:f)]) /
# sum(x[length(x)-f-(0:t)])
#aaply(y, c(1,3:6), function(x) maply(data.frame(i=(seq(length(x)-(t+f)))), function(i) fn(x[i:(i+t+f-1)])))

oms=FLStocks("Constant"=burnin[[2]],"No Demand"=nodemand[[2]],"Kicked"=kick[[2]])
om=FLQuants(llply(oms,function(x) biomass(x)[,,,3][,ac(2018:2024),,"3"]))
mp=FLQuants(llply(om, function(x) hDevs[,dimnames(x)$year]%*%x))

dat=maply(data.frame(year=2010:2030), 
          function(year) #apply(window(x,start=year-1,end=year),c(1,3:6),mean)%/%
                          #apply(window(x,start=year-4,end=year-2),c(1,3:6),mean))
            yearMeans(window(om[[1]], start=year-1,end=year)) /
              yearMeans(window(om[[1]], start=year-3,end=year-1)))
dat=as.FLQuant(transmute(melt(dat),year=year,iter=iter,data=value))

dt2=maply(data.frame(year=2010:2030), 
          function(year) #apply(window(y,start=year-1,end=year),c(1,3:6),mean)%/%
                         #  apply(window(y,start=year-4,end=year-2),c(1,3:6),mean))
            yearMeans(window(mp[[1]], start=year-1,end=year)) /
              yearMeans(window(mp[[1]], start=year-3,end=year-1)))
dt2=as.FLQuant(transmute(melt(dt2),year=year,iter=iter,data=value))


plot(FLQuants("OM"=dat,"Index"=dt2))
```

**Figure `r iFig=iFig+1; iFig`** $2/3$


```{r}
rtn=FLCore:::roc(dat[,ac(2019:2025)]>1,dt2[,ac(2019:2025)])

ggplot(rtn)+  
  geom_line(aes(FPR,TPR))+geom_abline(aes(intercept=0,slope=1),col="red")
```


**Figure `r iFig=iFig+1; iFig`** ROC Curve

## One over Two rule

```{r}
#fn=function(x,f=2,t=3) 
# sum(x[length(x)-(0:f)]) /
# sum(x[length(x)-f-(0:t)])
#aaply(y, c(1,3:6), function(x) maply(data.frame(i=(seq(length(x)-(t+f)))), function(i) fn(x[i:(i+t+f-1)])))

dat=maply(data.frame(year=2010:2030), 
          function(year) #apply(window(x,start=year-1,end=year),c(1,3:6),mean)%/%
                          #apply(window(x,start=year-4,end=year-2),c(1,3:6),mean))
            yearMeans(window(x, start=year,end=year)) /
              yearMeans(window(x, start=year-2,end=year-1)))
dat=as.FLQuant(transmute(melt(dat),year=year,iter=iter,data=value))  

dt2=maply(data.frame(year=2010:2030), 
          function(year) #apply(window(y,start=year-1,end=year),c(1,3:6),mean)%/%
                         #  apply(window(y,start=year-4,end=year-2),c(1,3:6),mean))
            yearMeans(window(y, start=year,end=year)) /
              yearMeans(window(y, start=year-2,end=year-1)))
dt2=as.FLQuant(transmute(melt(dt2),year=year,iter=iter,data=value))

plot(FLQuants("OM"=dat,"Index"=dt2))
```

**Figure `r iFig=iFig+1; iFig`** $1/2$


```{r}
rtn=FLCore:::roc(dat[,ac(2019:2025)]>1,dt2[,ac(2019:2025)])  
ggplot(rtn)+ 
  geom_line(aes(FPR,TPR))+geom_abline(aes(intercept=0,slope=1),col="red")
```


**Figure `r iFig=iFig+1; iFig`** ROC Curve

### Slope

```{r}
dat=maply(data.frame(year=2015:2025), function(year) aaply(window(x,start=year,end=year+4), c(1,3:6), function(idx) {
    dat=data.frame(x=names(idx),y=log(c(idx)/mean(idx)))  
    lm(y~x,data=dat)$coefficients[2]}))
dat=as.FLQuant(as.data.frame((FLPar(dat))))

dt2=maply(data.frame(year=2015:2025), function(year) aaply(window(y,start=year,end=year+4), c(1,3:6), function(idx) {
    dat=data.frame(x=names(idx),y=log(c(idx)/mean(idx)))
    lm(y~x,data=dat)$coefficients[2]}))
dt2=as.FLQuant(as.data.frame((FLPar(dt2))))

plot(FLQuants("OM"=dat,"Index"=dt2))
```

**Figure `r iFig=iFig+1; iFig`** Slope for 5 years


```{r}
rtn=FLCore:::roc(dat>0,dt2)  
ggplot(rtn)+
  geom_line(aes(FPR,TPR))+geom_abline(aes(intercept=0,slope=1),col="red")
```

**Figure `r iFig=iFig+1; iFig`** ROC Curve


```{r}
dat=maply(data.frame(year=2015:2025), function(year) aaply(window(x,start=year,end=year+2), c(1,3:6), function(idx) {
    dat=data.frame(x=names(idx),y=log(c(idx)/mean(idx)))
    lm(y~x,data=dat)$coefficients[2]}))
dat=as.FLQuant(as.data.frame((FLPar(dat))))  

dt2=maply(data.frame(year=2015:2025), function(year) aaply(window(y,start=year,end=year+2), c(1,3:6), function(idx) {
    dat=data.frame(x=names(idx),y=log(c(idx)/mean(idx)))
    lm(y~x,data=dat)$coefficients[2]}))
dt2=as.FLQuant(as.data.frame((FLPar(dt2))))

plot(FLQuants("OM"=dat,"Index"=dt2))
```

**Figure `r iFig=iFig+1; iFig`** Slope for 3 years


```{r}
rtn=FLCore:::roc(dat>0,dt2)   
names(rtn)=gsub(" ","_",names(rtn))
ggplot(rtn)+
  geom_line(aes(FPR,TPR))+geom_abline(aes(intercept=0,slope=1),col="red")
```

**Figure `r iFig=iFig+1; iFig`** ROC Curve


```{r,ssb-plot}
object=FLStocks("SQ"=burnin[[3]],"Predators"=burnin[[4]],"Kick"=kick[[4]],"No_demand"=nodemand[[4]])
mCtc=FLQuants(llply(object,function(x) apply(stock.wt(x)%*%m(x)/(z(x))%*%(1-exp(-z(x))),c(2,6),sum)))
mSSB=FLQuants(llply(object,function(x) ssb(x)[,,,2]))

plot(mSSB)+facet_grid(qname~.)
```

**Figure `r iFig=iFig+1; iFig`** SSB


```{r}
plot(mCtc)+facet_grid(qname~.)
```

**Figure `r iFig=iFig+1; iFig`** Predator share


```{r}
ggboxplot(rbind(cbind("Quantity"="SSB",     ldply(mSSB, function(x) as.data.frame(x[,"2040"],drop=T))),
                cbind("Quantity"="Predator",ldply(mCtc, function(x) as.data.frame(x[,"2040"],drop=T)))),
      x=".id",y="data")+
  facet_grid("Quantity~.",scale="free")
```

```{r}
load("../data/om/omScenarios.RData")

oms=list("Status Quo"=burnin,
         "Kicked"    =kick,
         "No_Demand" =nodemand,
         "F Trend"   =ftrend)

#### OM #########################################################################
om=mdply(data.frame(OM=names(oms)), function(OM) ldply(oms[[OM]], function(x) 
                      model.frame(FLQuants(SSB    =ssb(    x)[,,,2],
                                           Biomass=biomass(x)[,,,3],
                                           Fbar   =apply(fbar(   x),c(2,6),sum),
                                           Catch  =apply(catch(x),c(2,6),sum)),drop=T)))
      
#### MP #########################################################################
## LBIs
lbi=mdply(data.frame(OM=names(oms)), function(OM) ldply(oms[[OM]], function(x) 
                      as.data.frame(indicators.len(x[,,,3], indicators=c('lbar'), params=pars[,1], metric='catch.n')[[1]],drop=T)))

## Index
hDevs=    rlnoise(100,stock(oms[[1]][[1]])[,,,3,,1]%=%0,0.3,0)%*%
     qmin(rlnoise(100,stock(oms[[1]][[1]])[,,,3,,1]%=%0,0.5,0.6),1)

idx=mdply(data.frame(OM=names(oms)), function(OM) ldply(oms[[OM]], function(x) 
                      as.data.frame(biomass(x)[,,,3]%*%hDevs[,dimnames(biomass(x)[,,,3])$year],drop=T)))

smry=cbind(om,lbi=lbi[,"data"],idx=idx[,"data"])
```


```{r}
ggplot(smry)+ 
  geom_point(aes(Biomass,idx))+
  facet_grid(OM~.id)
```

# ROC & TSS

## LBIs
```{r}
load("../data/results/smry.RData")

lb=ddply(subset(smry,year%in%2021:2025),.(year,.id), with, roc(Fbar/(targetF*4),1/lbi))

ggplot(transform(lb,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock","F Trend"))))+
  geom_path(aes(FPR,TPR,col=year))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")
```

**Figure `r iFig=iFig+1; iFig`** ROC curves for LBIs


```{r}
val=ddply(smry,.(.id,OM,iter), with, mdply(data.frame(yr=2021:2030), function(yr)
                      data.frame(value=mean(lbi[year%in%(yr+0)]))))
ref=ddply(subset(smry,year%in%2010:2011), .(OM,.id,iter), with, data.frame(ref=mean(lbi)))
lb =merge(val,ref)

lb=ddply(lb,.(yr,.id), with, roc(value/ref,value))
```

```{r}
ggplot(transform(lb,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=yr))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")

tss=ddply(lb, .(.id,yr), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})
```

```{r}
ggplot(tss)+
  geom_line( aes(yr,indicator,col=.id))+
  geom_point(aes(yr,indicator,col=.id))
```

```{r}
ggplot(tss)+
  geom_line( aes(yr,TSS,col=.id))+
  geom_point(aes(yr,TSS,col=.id))
```

**Figure `r iFig=iFig+1; iFig`** TSS for $1/2$


```{r}
val=ddply(smry,.(.id,OM,iter), with, mdply(data.frame(yr=2021:2030), function(yr) data.frame(value=mean(lbi[year%in%(yr+0+1)]))))
ref=ddply(subset(smry,year%in%2010:2012), .(OM,.id,iter), with, data.frame(ref=mean(lbi)))
lb =merge(val,ref)

lb=ddply(lb,.(yr,.id), with, roc(value/ref,value))

ggplot(transform(lb,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=yr))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")

tss=ddply(lb, .(.id,yr), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})

ggplot(tss)+
  geom_line( aes(yr,indicator,col=.id))+
  geom_point(aes(yr,indicator,col=.id))

ggplot(tss)+
  geom_line( aes(yr,TSS,col=.id))+
  geom_point(aes(yr,TSS,col=.id))
```

**Figure `r iFig=iFig+1; iFig`** TSS for $1/2$


## Harvest Rate
```{r, hr2}
smry$hr=smry$Catch/smry$Biomass

hr=ddply(subset(smry,year%in%2021:2025),.(year,.id), with, roc(Fbar/(targetF*4),hr))

ggplot(transform(hr,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=year))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")
```

**Figure `r iFig=iFig+1; iFig`** ROC curves for HR


```{r, hr2-plot}
val=ddply(smry,.(.id,OM,iter), with, mdply(data.frame(yr=2021:2030), function(yr) data.frame(value=mean(hr[year%in%(yr+0)]))))
ref=ddply(subset(smry,year%in%2010:2011), .(OM,.id,iter), with, data.frame(ref=mean(hr)))
hr =merge(val,ref)

hr=ddply(hr,.(yr,.id), with, roc(value/ref,value))

ggplot(transform(hr,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=yr))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")

tss=ddply(hr, .(.id,yr), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})

ggplot(tss)+
  geom_line( aes(yr,indicator,col=.id))+
  geom_point(aes(yr,indicator,col=.id))

ggplot(tss)+
  geom_line( aes(yr,TSS,col=.id))+
  geom_point(aes(yr,TSS,col=.id))
```

**Figure `r iFig=iFig+1; iFig`** TSS for $1/2$


```{r}
val=ddply(smry,.(.id,OM,iter), with, mdply(data.frame(yr=2021:2030), function(yr) data.frame(value=mean(hr[year%in%(yr+0+1)]))))
ref=ddply(subset(smry,year%in%2010:2012), .(OM,.id,iter), with, data.frame(ref=mean(hr)))
hr =merge(val,ref)

hr=ddply(hr,.(yr,.id), with, roc(value/ref,value))

ggplot(transform(hr,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=yr))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")

tss=ddply(hr, .(.id,yr), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})

ggplot(tss)+
  geom_line( aes(yr,indicator,col=.id))+
  geom_point(aes(yr,indicator,col=.id))

ggplot(tss)+
  geom_line( aes(yr,TSS,col=.id))+
  geom_point(aes(yr,TSS,col=.id))
```

**Figure `r iFig=iFig+1; iFig`** TSS for $1/2$

## Index
```{r}
idx=ddply(subset(smry,year%in%2021:2025),.(year,.id), with, roc(Biomass/mean(Biomass),idx))

ggplot(transform(idx,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=year))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")
```

**Figure `r iFig=iFig+1; iFig`** ROC curves for HR


```{r}
ggplot(smry)+
  geom_point(aes(Biomass,idx))+
  facet_grid(OM~.id)
```

```{r}
val=ddply(smry,.(.id,OM,iter), with, mdply(data.frame(yr=2021:2030), function(yr)
     data.frame(value=mean(idx[year%in%(yr+0)]))))
ref=ddply(subset(smry,year%in%2010:2011), .(OM,.id,iter), with, data.frame(ref=mean(idx)))
idx =merge(val,ref) 

idx=ddply(idx,.(yr,.id), with, roc(value/ref,value))

ggplot(transform(idx,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=yr))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")

tss=ddply(idx, .(.id,yr), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})

ggplot(tss)+
  geom_line( aes(yr,indicator,col=.id))+
  geom_point(aes(yr,indicator,col=.id))

ggplot(tss)+
  geom_line( aes(yr,TSS,col=.id))+
  geom_point(aes(yr,TSS,col=.id))
```

**Figure `r iFig=iFig+1; iFig`** TSS for $1/2$


```{r}
val=ddply(smry,.(.id,OM,iter), with, mdply(data.frame(yr=2021:2030), function(yr) data.frame(value=mean(idx[year%in%(yr+0+1)]))))
ref=ddply(subset(smry,year%in%2010:2012), .(OM,.id,iter), with, data.frame(ref=mean(idx)))
idx =merge(val,ref)

idx=ddply(idx,.(yr,.id), with, roc(value/ref,value))

ggplot(transform(idx,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR,col=yr))+
  facet_wrap(~.id)+
  geom_abline(aes(slope=1,intercept=0),col="red")

tss=ddply(idx, .(.id,yr), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})

ggplot(tss)+
  geom_line( aes(yr,indicator,col=.id))+
  geom_point(aes(yr,indicator,col=.id))

ggplot(tss)+
  geom_line( aes(yr,TSS,col=.id))+
  geom_point(aes(yr,TSS,col=.id))
```

**Figure `r iFig=iFig+1; iFig`** TSS for $1/2$

# Slope
```{r}
dat=ddply(smry,.(OM,.id,iter), with, mdply(data.frame(yr=2021:2030), 
                                                    function(yr) {
                                                      idx=log(idx/mean(idx)) 
                                                      dat=data.frame(year=year[year%in%(yr+0:2)],idx=idx[   year%in%(yr+0:2)])
                                                      
                                                      bio=log(Biomass/mean(Biomass)) 
                                                      dt2=data.frame(year=year[year%in%(yr+0:2)],om=bio[year%in%(yr+0:2)])
                                                  
                                                      data.frame(idx=coefficients(lm(idx~year,data=dat))[2],
                                                                 om =coefficients(lm(om ~year,data=dt2))[2])
                                                      }))

ggplot(dat)+
  geom_point(aes(om,idx,col=OM))+
  facet_wrap(~.id)+
  geom_abline(aes(intercept=0,slope=1,col=OM))

slope=ddply(dat,.(.id), with, roc(om+1,idx))

ggplot(transform(slope,.id=factor(.id,levels=c("Rec","M Annual","M Cohort","M Shock"))))+
  geom_path(aes(FPR,TPR))+
  geom_abline(aes(slope=1,intercept=0),col="red")+
  facet_wrap(~.id)

tss=ddply(lb, .(.id), with, { flag=TSS==max(TSS); data.frame(TSS=TSS[flag],indicator=indicator[flag])})

ggplot(tss)+
  geom_line( aes(.id,indicator))+
  geom_point(aes(.id,indicator))

ggplot(tss)+
  geom_line( aes(.id,TSS))+
  geom_point(aes(.id,TSS))
```

Biomass safeguards can be based on the following Index (I) reference levels:

    Imin (= historical minimum value),
    Itrig (= 1.4 Imin) 
    Istat (=geometricmean(Ihist) exp(-1.645 sd(log(Ihist))

```{r}
lmin=ddply(subset(smry,year%in%1980:2020), .(.id,iter), with, 
      data.frame(lmin =min(idx),
                 ltrig=min(idx)*1.4,
                 lstat=exp(mean(log(idx)))*exp(-1.645*var(log(idx))^0.5)))

ggboxplot(data=melt(lmin,id=c(".id","iter")),y="value",x=".id")+
  facet_grid(.~variable)
```


```{r, save}
save(hDevs,                            file="../data/om/dev-idx.RData")
save(smry,                             file="../data/results/smry.RData")
save(burnin,kick,nodemand,eq4,pars,sr4,file="../data/results/skill.RData")
```
